---
title: "Permutación y Correlación Ejemplos"
author: "F.A. Barrios"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
knitr::opts_chunk$set(cache = TRUE)

## Global options
options(max.print="80")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=80)
```

```{r}
library(tidyverse)
```

Ejemplos resueltos de Permutación y Correlación, tomados de los problemas del capítulo 9 del Daniel.  

**Ejemplo** Dean Parmalee wished to know if the year-end grades assigned to Wright State University Medical School students are predictive of their second-year board scores. The following table shows, for 89 students, the year-end score (AVG, in percent of 100) and the score on the second-year medical board examination (BOARD). Table in REV_C09_17.csv. Perform a complete regression analysis with AVG as the independent variable. Let α = .05 for all tests.  
*Respuesta* Describir los datos. Correr el modelo lineal para Board ~ AVG con $\alpha = 0.05$.  

```{r}
Scores <- read_csv("DataSets/ch09_all/REV_C09_17.csv", show_col_types = FALSE)
Scores

ggplot(Scores, aes(x = AVG, y = BOARD)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(BOARD)), linetype = "dashed", color = "red") +
  geom_vline(aes(xintercept = mean(AVG)), linetype = "dashed", color = "red")

Scores_lm <- lm(BOARD ~ AVG, data = Scores)
summary(Scores_lm)

ggplot(Scores, aes(x = AVG, y = BOARD)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(BOARD)), linetype = "dashed", color = "red") +
  geom_vline(aes(xintercept = mean(AVG)), linetype = "dashed", color = "red") +
  geom_abline(aes(intercept = Scores_lm$coefficients[1], slope = Scores_lm$coefficients[2]), color = "blue")

plot(Scores_lm)
```
